\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,booktabs,graphicx,hyperref,natbib}
\title{From CLV Ranking to Uplift-Aware Targeting:\\
Cross-Context Evidence from Online Retail, Grocery, and Supermarket Campaign Data}
\author{Group 11}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Customer lifetime value (CLV) models identify \emph{who is valuable}; uplift models identify
\emph{who is responsive} to marketing. We connect the two across three retail contexts. On
UK online retail and US grocery data we show an explainable Hurdle--Semantic CLV framework
transfers across contexts but for \emph{different reasons}: in zero-inflated e-commerce the
return-incidence stage dominates, while in habitual grocery shopping temporal-sequence signal
dominates. On the X5 RetailHero campaign data (randomised treatment/control), we show that
ranking customers by predicted response is \emph{worse than random} for incremental targeting,
whereas uplift learners recover persuadable customers. Finally, a \emph{value-adjusted uplift}
score ($\hat\tau_i\cdot\hat V_i$) yields the highest simulated profit under tight budgets,
though we report transparently that the dollar differences are directional rather than
statistically significant. CLV tells us who is valuable; uplift tells us who is responsive;
value-adjusted uplift tells us who is worth targeting.
\end{abstract}

\section{Introduction}
CLV studies typically rank high-value customers, while uplift studies rank campaign-responsive
customers. Yet a high-CLV customer may purchase regardless of intervention, and a mid-value but
high-uplift customer may be the true ``persuadable''. Limited work systematically connects
\emph{explainable CLV ranking} with \emph{uplift-aware targeting} across multiple retail contexts.
We study four questions:
\begin{itemize}
  \item[\textbf{RQ1}] How robust is the explainable Hurdle--Semantic CLV framework across UK
  online retail and US grocery contexts?
  \item[\textbf{RQ2}] Do semantic product profiles and behavioural sequence features add
  incremental value beyond RFM for CLV ranking, calibration, and interpretation?
  \item[\textbf{RQ3}] On campaign data with treatment/control labels, does uplift-aware targeting
  outperform CLV-only, RFM-only, and response-model targeting?
  \item[\textbf{RQ4}] Can combining predicted value with uplift improve simulated campaign
  efficiency, and what transferable principles follow for budget-constrained markets?
\end{itemize}

\section{Related Work}
Zero-inflated / heavy-tailed CLV motivates Hurdle and ZILN decompositions \citep{cragg1971hurdle,
wang2019deep}; sequence features improve CLV \citep{bauer2021improved}; customer embeddings give
semantic representations \citep{chamberlain2017customer}. Conformalized quantile regression gives
valid prediction intervals \citep{romano2019cqr}. Uplift modelling estimates incremental treatment
effect \citep{gutierrez2017uplift}; meta-learners and causal forests estimate heterogeneous
effects \citep{kunzel2019metalearners,wager2018causalforest}; causal ML supports campaign
targeting \citep{gubela2024multiple,fader2005rfm}.

\section{Data and Problem Setup}
\textbf{Online Retail II} (UK e-commerce, Phase~1 baseline, read-only) and \textbf{Dunnhumby
Complete Journey} (US grocery, external CLV validation) provide CLV ranking; \textbf{X5
RetailHero} (randomised SMS campaign) provides treatment/control for uplift. We deliberately do
\emph{not} use Dunnhumby coupon redemption as a causal treatment (it is a post-campaign outcome,
not clean assignment). CLV target $Y_i=\sum_{t\in\text{horizon}}\text{Revenue}_{i,t}$.

\section{Methodology}
\textbf{Module A --- CLV ranking (Hurdle).} $\widehat{CLV}_i=p_i\cdot m_i$ with
$p_i=P(Y_i>0\mid X_i)$ and $m_i=E[Y_i\mid Y_i>0,X_i]$, using a lognormal bias correction.
\textbf{Module B --- Semantic profiling.} Product co-purchase PPMI $\to$ truncated SVD
embeddings; customer taste vector = revenue-weighted mean of bought-product embeddings.
\textbf{Module C --- Uplift.} $\tau_i=E[Y_i(1)-Y_i(0)\mid X_i]$ via S/T/X-Learners and class
transformation. \textbf{Module D --- Value-adjusted uplift.}
$\text{Score}^{\text{VU}}_i=\hat\tau_i\cdot\hat V_i$, with $\hat V_i$ a value proxy from
historical spend (\emph{not} actual future CLV).

\section{Experiments and Results}
We use a stratified 80/20 split (CLV) and 70/30 split (uplift), seed 42. Metrics: Normalized
Gini (NG), Revenue Capture@10\% (RC@10), Precision@10\%, decile calibration; Qini AUC, AUUC,
uplift@K; CQR coverage; bootstrap CIs.

\input{tables_auto}

\textbf{RQ1/RQ2.} Table~\ref{tab:clv_summary}: Hurdle is best on e-commerce (NG$=0.91$) but ties
a linear model on grocery (NG$=0.85$); extra features beat RFM significantly in both contexts
(Table~\ref{tab:robust}), with a larger gain in grocery. SHAP shows monetary drives spend in
e-commerce while recent-sequence spend drives it in grocery; CQR attains valid coverage
(90\%$\to$90.2\%/89.0\%).

\textbf{RQ3.} Table~\ref{tab:uplift}: the response model ranks \emph{below random} (Qini
$-0.007$, CI entirely negative), while uplift learners reach $\approx 11\%$ uplift in the top
decile --- ``likely to buy'' $\neq$ ``persuadable''.

\textbf{RQ4.} Table~\ref{tab:policy}: targeting by value alone can lose money at small budgets;
uplift-only maximises incremental conversions; value-adjusted uplift gives the highest simulated
profit at tight budgets. Bootstrap CIs (Table~\ref{tab:robust}) confirm the ranking claims are
significant, but the \emph{dollar} profit differences are directional, not significant.

\section{Discussion: Transferable Insights}
(1) Do not target on CLV alone --- it selects ``sure buyers''. (2) Do not target on response
probability alone --- it selects low-incremental customers. (3) Value-adjusted uplift fits
budget-constrained settings. (4) Explainability clarifies \emph{why} a customer is selected.
(5) Uncertainty matters when market data are unstable. These provide transferable design
principles for budget-constrained emerging retail markets.

\section{Limitations}
Single train/test split per context (mitigated by bootstrap); X5 value proxy is historical, not
future, spend; profit differences between policies are not statistically significant; uplift
absolute Qini is modest (the campaign signal itself is small, $\approx 3.3$pp).

\section{Conclusion}
Across three retail contexts we connect explainable CLV ranking with uplift-aware targeting. The
framework transfers but its mechanism is context-dependent, and value-adjusted uplift is the most
promising targeting rule under budget constraints.

\bibliographystyle{plainnat}
\bibliography{references}
\end{document}
